Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface

### 2022-04-11 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
Add MR edits

---
title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface"
author:
  - name: Brett Longworth
    email: blongworth@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
    correspondingauthor: true
  - name: Josh Burton
    email: jburton@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
  - name: Simon Pendleton
    email: simon.pendleton@plymouth.edu
    affiliation: Plymouth State University
  - name: Sydney Moser
    email: smose007@uottawa.ca
    affiliation: University of Ottawa
  - name: Mark Roberts
    email: mroberts@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
  - name: Mark Kurz
    email: mkurz@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
address:
  - code: Woods Hole Oceanographic Institution
    address: Department of Geology and Geophysics, 360 Woods Hole Rd., Woods Hole, MA, 02540
  - code: Plymouth State University
    address: Department, Street, City, State, Zip
  - code: University of Ottawa
    address: Department, Street, City, State, Zip
footnote:
abstract: |
  We have developed a method for measuring radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method uses helium as a carrier gas to displace CO~2~ from sample vials to an open split, where a glass capillary samples the mixture for delivery directly to the HGIS. This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement. Samples larger than 8 mg carbonate can be measured. Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4 - 1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1$\sigma$ uncertainty is about 1 %. We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
keywords: 
  - radiocarbon
  - carbonate
  - hybrid gas ion source
journal: "Nuclear Instruments and Methods in Physics Research Section B"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: HGIS_paper.bib
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
csl: elsevier-without-titles.csl
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

@@

Most interfaces to gas-accepting sputter ion sources are a closed inlet design [@Xu2007; @Uhl2007; @ruff_gas_2007; @ruff_-line_2010; @Fahrni2013; @wacker_versatile_2013; @gottschalk_radiocarbon_2018]. A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the inlet volume and the length and diameter of the capillary. Pressure is adjusted by changing either the inlet volume or the amount of gas in the inlet. We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere, often called an open split. This defines the delivery pressure as 1 ATM, and therefore flow is controlled by the capillary parameters and is not easily adjustable during measurement. The flow of CO~2~ to the source may be varied by controlling the ratio of CO~2~ to helium in the sample stream. Changing the ratio of CO~2~ to helium changes the partial pressure of CO~2~ and therefore the carbon flow rate to the source.

@@

The open split interface (Figure \ref{fig1}) supplies a helium - CO~2~ mixture to a split where a capillary delivers gas to the source and the rest is vented. Flowing helium is used to displace sample CO~2~ from the sample vial to the delivery capillary in the split. Our open-split interface is constructed of simple, off-the-shelf parts. The core of the design is a dual needle and split housed in a VICI cross fitting. Displacement helium flows in through one side of the cross and into the sample vial through 1/16 in. stainless steel tubing, which forms the outer flow path of the concentric dual needle.

The mixture of sample CO~2~ and displacement helium from the vial flows up a 21 gauge (0.819 mm) needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle. The open split is a VICI Luer adapter (part ZBUMLPK) with 5cm of 1/16 in. PFTE tubing attached. The split is connected to the top of the inner needle, and the source delivery capillary is inserted into the split. The split has fairly low volume (< 1 mL). The system was operated manually during development, but may be mounted on the arm of a gasbench-style autosampler to enable automated operation.

@@

The helium displacement flow rate was set to 250 $\mu$l min^-1^ for initial tests, and this was reduced to 100 $\mu$l min^-1^ for later work, including most reference standard comparisons and the core comparison. The flow rate from the open split into the source was defined by 150 cm of 50 $\mu$m fused silica capillary, split into two sections by a zero-volume bulkhead connector at the source high-voltage cage. The section inside the cage was jacketed in 1/16 in. PFTE tubing to help protect against spark damage. The total flow rate of gas to the source varied due to the viscosity difference between helium and CO~2~ and helium, and was estimated to be between 30 $\mu$l min^-1^ for pure helium and 40 $\mu$l min^-1^ for pure CO~2~ using the Hagen-Poiseuille equation.

```{r fig1, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig1}Schematic of the open split inlet. Two valco 2-way valves provide the ability to admit helium and live or dead tank standards. The double needle is housed in a VICI cross and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig1_HGIS_open_split_drawing.pdf"))
```

@@

Carbonate samples were weighed into 7 ml septa-sealed Vacutainer vials (Labco 768W). Sealed vials were evacuated, then 1 ml 85% phosphoric acid was added. The samples were allowed to react at 60 C for 6 h. Sample pressure was adjusted to 1 atm by allowing the sample to equilibrate with a helium-filled syringe connected through the vial septum. Evacuation, acid addition, and heated reaction were performed by the automated gas handling system developed by Roberts et al. [@Roberts2013].

@@

Standards were prepared by evacuating vials and then filling with a modern standard CO~2~ using the automated gas handling system. Standards were adjusted to the same size as other samples by extracting an appropriate volume of CO~2~ from the full vial and replacing with helium using a syringe. Process blank samples were prepared from powdered C-1 or TIRI-F, acidified in vials using the method above.

@@

Samples were measured on a modified 3MV Tandetron AMS system [@Longworth2015]. The displacement helium was allowed to flow constantly during analysis, and the sample needle was inserted into a helium-filled vial between samples. Measurement of a sample proceeded as follows: a clean sputter target cathode was rotated into position and the target was pre-cleaned by sputtering for 5 minutes or until cathode ^12^C^-^ current fell below 100 nA. The delivery needle was then transferred to the sample vial, which started sample delivery to the split and source. Currents were allowed to stabilize for 30 s, at which point the data acquisition was started. Data was acquired in 10 blocks of 90 s each, for 15 min total acquisition time. Normalizing standards, blanks, and secondary standards were measured identically to unknowns. Total time for a measurement was 20-25 min including target pre-cleaning and sample changes.

@@

Measurements were corrected for process and measurement blank using a linear correction that assumes that the corrected fraction modern of blank samples should be zero, samples with the fraction modern of the standard need no correction, and applies a proportional correction to samples of intermediate fraction modern [@donahue_isotope-ratio_1990; @roberts_14c_2019]. The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank. The errors in the blanks and standards were propagated with errors of the samples to produce the final error estimate for the sample. This correction does not account for the effect of a blank with constant mass or current on samples of varying mass or current. For all measurements used to assess the performance of the method and for the Haiti core comparison, sample, standard, and blank masses were kept constant. Sample current did vary, but we believe the linear correction described above produces the most accurate results within this range. We discuss this constant contaminant blank in the results below. Data were reduced using a package written in R [@r_core_team_r_2022] which is published on github (<https://github.com/blongworth/hgis>).

@@

```{r fig2, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig2}Performance of the NOSAMS MCSNICS hybrid gas ion source. (a) Measured $^{12}C^-$ ion current during displacement of a pure $CO_2$ vial sample with helium. (b) Modeled flows of helium and $CO_2$ to the ion source during the same dilution experiment. (c) Measured ion current for modeled $CO_2$ flow. The stability of the source over a range of flow produces stable current for up to 40 min of measurement time.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.pdf"))
```

@@

$$C(t) = e^{-\frac{r}{V}t}$$

Where $C(t)$ is the fraction CO~2~, $r$ is the rate of helium displacement flow, and $V$ is the initial volume of the CO~2~-filled vial. Figure 2c shows the resulting relationship of sample ion current to CO~2~ flow rate. The most important feature of the current-flow relationship is the wide range of flows under which stable ion current is measured, which in turn produces relatively stable currents and ratios despite a changing carbon flow rate during measurement. Several studies have measured peak currents over a narrower range of flows [@Fahrni2013; @broek_optimization_nodate] at high helium dilution rates (> 10:1), while Xu et. al [@Xu2007] shows a somewhat broader peak with pure CO~2~. Broek et. al [@broek_optimization_nodate] found peak broadening with decreasing dilution ratio. While lower rates of dilution used in our HGIS method may decrease current slightly, the observed peak broadening allows stable measurement over the range of CO~2~ flows produced by the open split.

Work with currents and flows (above and Figure 2) predicts that optimal currents should be produced until less than 2 ml CO~2~ (equivalent to 0.8 mg C) remains in the sample vial. Tests of a range of sample sizes support this estimate and show that no reduction in measured current is seen in samples larger than 1 mg C which corresponds to 8 mg CaCO~3~ sample material. To ensure ample sample for optimal measurements, we chose 15 mg as a target sample mass. To ensure that samples started at or slightly above atmospheric pressure, we injected clean helium with a syringe to equilibrate pressure in the vials when the reaction with acid was complete.

@@

The HGIS method produced higher ratios for radiocarbon dead samples than the graphite method. The average normalized ratio for radiocarbon-free reference materials IAEA C1 marble and TIRI F spar calcite was `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`). We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank, and implemented a blank correction for the method.

@@

```{r fig3, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig3}Assessment of blanks. (a) Current dependent blank model and data. $F^{14}C$ and current of measurements from dilution tests of live and dead gas are used to generate linear fits which should cross at the $F^{14}C$ and current of the constant current blank. Direct determination of $F^{14}C$ and current of the blank are also shown. (b) Fraction modern of large, radiocarbon dead samples compared to sample current. Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig3_blanks.pdf"))
```

@@

$$ R_{samp} = R_{meas} - \frac{R_{blank}(R_{std}-R_{meas})}{R_{std}} $$

@@

## Reference materials

@@

Precision and accuracy of the method were assessed by measuring carbonate reference standards for comparison with consensus values from graphite measurements. We measured TIRI I, a roughly quarter-modern turbidite carbonate, IAEA C2, a half modern turbidite carbonate, and NOSAMS 2, a modern internal shell carbonate standard from a bivalve mollusk REF. Summary data for these samples are given in Table 1, and measurement data for all samples is in supplementary information. Measurement, data reduction, and blank correction was performed as described above.

@@

```{r fig4, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig4}Agreement of standards with graphite consensus values. Difference of HGIS measurements from consensus for individual samples are plotted against their expected value. Expected values are spread to avoid overlapping points. Error bars represent one sigma sample uncertainty. The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.pdf"))
```

>>>>

---
title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface"
author:
  - name: Brett Longworth
    email: blongworth@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
    correspondingauthor: true
  - name: Josh Burton
    email: jburton@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
  - name: Simon Pendleton
    email: simon.pendleton@plymouth.edu
    affiliation: Plymouth State University
  - name: Sydney Moser
    email: smose007@uottawa.ca
    affiliation: University of Ottawa
  - name: Mark Roberts
    email: mroberts@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
  - name: Mark Kurz
    email: mkurz@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
address:
  - code: Woods Hole Oceanographic Institution
    address: Department of Geology and Geophysics, 360 Woods Hole Rd., Woods Hole, MA, 02540
  - code: Plymouth State University
    address: Environmental Science and Policy, 17 High St., Plymouth, MA, 03264
  - code: University of Ottawa
    address: Geography, Environment and Geomatics, 550 Cumberland St., Ottawa, ON, Canada, K1N 6N5
abstract: |
  We have developed a method for measuring radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method uses helium as a carrier gas to displace CO~2~ from sample vials to an open split, where a glass capillary samples the mixture for delivery directly to the HGIS. This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement. Samples larger than 8 mg carbonate can be measured. Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4 - 1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1$\sigma$ uncertainty is about 1 %. We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
keywords: 
  - radiocarbon
  - carbonate
  - hybrid gas ion source
journal: "Nuclear Instruments and Methods in Physics Research Section B"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: ../HGIS_paper.bib
linenumbers: false
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
csl: elsevier-without-titles.csl
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

@@

Most interfaces to gas-accepting sputter ion sources are a closed inlet design [@Xu2007; @Uhl2007; @ruff_gas_2007; @ruff_-line_2010; @Fahrni2013; @wacker_versatile_2013; @gottschalk_radiocarbon_2018]. A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the inlet volume and the length and diameter of the capillary. Pressure is adjusted by changing either the inlet volume or the amount of gas in the inlet. We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere, often called an open split. This defines the delivery pressure as 1 ATM. CO~2~ flow to the source can be controlled by changing the length or diameter of the capillary, or by changing the ratio of CO~2~ to He in the sample stream.

@@

The open split interface (Figure \ref{fig1}) supplies a helium - CO~2~ mixture to a split where a capillary delivers gas to the source and the rest is vented. Flowing helium is used to displace sample CO~2~ from the sample vial to the delivery capillary in the split. Our open-split interface is constructed of simple, off-the-shelf parts. The core of the design is a dual needle and split housed in a Valco tee fitting. Displacement helium flows in through the sidearm of the tee and into the sample vial through 1/16 in. stainless steel tubing, which forms the outer flow path of the concentric dual needle.

The mixture of sample CO~2~ and displacement helium from the vial flows up a 21 gauge (0.819 mm) needle inserted through the tee and stainless tubing which forms the inner flow path of the dual needle. The open split is a VICI Luer adapter (part ZBUMLPK) with 5cm of 1/16 in. PFTE tubing attached. The split is connected to the top of the inner needle, and the source delivery capillary is inserted into the split. The split has fairly low volume (< 1 mL). The system was operated manually during development, but may be mounted on the arm of a vial-sampling autosampler to enable automated operation.

@@

The helium displacement flow rate was set to 250 $\mu$l min^-1^ for initial tests, and this was reduced to 100 $\mu$l min^-1^ for later work, including most reference standard comparisons and the core comparison. The flow rate from the open split into the source was defined by 150 cm of 50 $\mu$m diameter fused silica capillary, split into two sections by a zero-volume bulkhead connector at the source high-voltage cage. The section inside the cage was jacketed in 1/16 in. PFTE tubing to help protect against spark damage. The total flow rate of gas to the source varied between 30 $\mu$l min^-1^ for pure helium and 40 $\mu$l min^-1^ for pure CO~2~ due to the viscosity difference between helium and CO~2~, estimated using the Hagen-Poiseuille equation.

```{r fig1, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig1}Schematic of the open split inlet. Two valco 2-way valves provide the ability to admit helium and live or dead tank standards. The double needle is housed in a tee fitting and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig1_HGIS_open_split_drawing.pdf"))
```

@@

Carbonate samples were weighed into 7 ml septa-sealed glass vials (Labco 768W). Sealed vials were evacuated, then 1 ml 85% phosphoric acid was added. The samples were allowed to react at 60 C for 6 h. Sample pressure was adjusted to 1 atm by allowing the sample to equilibrate with a helium-filled syringe connected through the vial septum. Evacuation, acid addition, and heated reaction were performed by the automated gas handling system developed by Roberts et al. [@Roberts2013].

@@

Normalizing standards were prepared by evacuating vials and then filling with a modern standard CO~2~ using the automated gas handling system. Standards were adjusted to the same size as other samples by extracting an appropriate volume of CO~2~ from the full vial and replacing with helium using a syringe. Process blank samples were prepared from powdered IAEA C1 or TIRI F, acidified in vials using the method above.

@@

Samples were measured on a modified 3MV Tandetron AMS system [@Longworth2015]. Measurement of a sample proceeded as follows: a clean sputter target cathode was rotated into position and the target was pre-cleaned by sputtering for 5 minutes or until cathode ^12^C^-^ current fell below 100 nA. The delivery needle was then transferred to the sample vial, which started gas delivery to the split and source. Currents were allowed to stabilize for 30 s, at which point the data acquisition was started. Data was acquired in 10 blocks of 90 s each, for 15 min total acquisition time. The displacement helium was allowed to flow constantly during cathode cleaning and analysis, and the sample needle was inserted into a helium-filled vial between samples. Normalizing standards, blanks, and secondary standards were measured identically to unknowns. Total time for a measurement was 20-25 min including target pre-cleaning and changing samples.

@@

Measurements were corrected for process and measurement blank using a linear correction that assumes that the corrected ratio of blank samples should be zero, samples with the same ratio as the standard need no correction, and samples of intermediate ratio need a proportional correction [@donahue_isotope-ratio_1990; @roberts_14c_2019]. The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank. The errors in the blanks and standards were propagated with errors of the samples to produce the final error estimate for the sample. This correction does not account for the effect of a blank with constant mass or current on samples of varying mass or current. For all measurements used to assess the precision and accuracy of the method, sample, standard, and blank masses were kept constant. Sample current did vary, but we believe the linear correction described above produces the most accurate results within this range. We discuss this constant contaminant blank in the results below. Data were reduced using a package written in R [@r_core_team_r_2022] which is published on github (<https://github.com/blongworth/hgis>).

@@

```{r fig2, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig2}Performance of the NOSAMS MCSNICS hybrid gas ion source. (a) Measured $^{12}C^-$ ion current during displacement of a pure $CO_2$ vial sample with helium. (b) Modeled flows of helium and $CO_2$ to the ion source during the same dilution experiment. (c) Measured ion current for modeled $CO_2$ flow. The stability of the source over a range of flow produces stable current for up to 40 min of measurement time.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.pdf"))
```

@@

\begin{equation}
\label{eq1}
C(t) = e^{-\frac{r}{V}t}
\end{equation}

Where $C(t)$ is the fraction CO~2~, $r$ is the rate of helium displacement flow, and $V$ is the initial volume of the CO~2~-filled vial. Figure 2c shows the resulting relationship of sample ion current to CO~2~ flow rate. The most important feature of the current-flow relationship is the wide range of CO~2~ flow rates over which stable ion current is measured, which in turn produces relatively stable currents and ratios despite a changing carbon flow rate during measurement. Several studies have measured peak currents over a narrower range of flows [@Fahrni2013; @broek_optimization_nodate] at high helium dilution rates (> 10:1), while Xu et. al [@Xu2007] shows a somewhat broader peak with pure CO~2~. Broek et. al [@broek_optimization_nodate] found that the range of flow rates with peak currents increased with decreasing dilution ratio. While lower rates of dilution used in our HGIS method may decrease current slightly, the observed peak broadening allows stable measurement over the range of CO~2~ flows produced by the open split.

Work with currents and flows (above and Figure 2) shows that stable currents are produced until less than 2 ml CO~2~ (0.8 mg C) remains in the sample vial. Tests of a range of sample sizes support this estimate and show that no reduction in measured current is seen in samples larger than 1 mg C which corresponds to 8 mg CaCO~3~ sample material. To ensure ample sample for optimal measurements, we chose 15 mg as a target sample mass. To ensure that samples started at or slightly above atmospheric pressure, we injected helium with a syringe to equilibrate pressure in the vials when the reaction with acid was complete.

@@

The HGIS method produced higher ratios for radiocarbon dead samples than the graphite method. The average uncorrected $F^{14}C$ for radiocarbon-free reference materials IAEA C1 marble and TIRI F spar calcite was `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`). We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank, and implemented a blank correction for the method.

@@

```{r fig3, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig3}Assessment of blanks. (a) Current dependent blank model and data. $F^{14}C$ and current of measurements from dilution tests of live and dead gas are used to generate linear fits which should cross at the $F^{14}C$ and current of the constant current blank. Direct determination of $F^{14}C$ and current of the blank are also shown. (b) Fraction modern of large, radiocarbon dead samples compared to sample current. Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig3_blanks.pdf"))
```

@@

\begin{equation}
\label{eq2}
R_{samp} = R_{meas} - \frac{R_{blank}(R_{std}-R_{meas})}{R_{std}}
\end{equation}

@@

## Standard reference materials

@@

Precision and accuracy of the method were assessed by measuring carbonate reference standards for comparison with consensus values from graphite measurements. We measured TIRI I, a roughly quarter-modern turbidite carbonate, IAEA C2, a half modern turbidite carbonate, and NOSAMS 2, a modern internal shell carbonate standard from a bivalve mollusk. Summary data for these samples are given in Table 1, and measurement data for all samples is in supplementary information. Measurement, data reduction, and blank correction was performed as described above.

@@

```{r fig4, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig4}Agreement of standards with graphite consensus values. Difference of HGIS measurements from consensus for individual samples are plotted against their expected value. Expected values are spread to avoid overlapping points. Error bars represent one sigma sample uncertainty. The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.pdf"))
```
